### 2020-10-07 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
comments in russian

@@

A note worthy approach to solving MCO problems is lexicographic optimization methods. Such methods involve ranking the criteria by the degree of their importance and using the obtained ranking for subsequent optimization \cite{c3}. When criteria are ranked by importance, an MCO problem can be solved via consecutive optimization of criteria in the order of their importance: the most important criterion is optimized first, followed by the optimization of the second criterion based on the analysis of the set of decisions, in which the first criterion reaches the best value. This sequence is simple and clear and allows using many efficient scalar optimization algorithms. Such a multistage sequence can be reduced to solving one scalar optimization problem by linear convolution of criteria \cite{c38}. At the same time, this approach excessively reduces the number of calculated efficient decisions, which limits the number of options for the decision-maker. The $\varepsilon$-constraint scalarization approach can be used to expand the set of efficient decisions that can be considered by the decision-maker. In this case, only the most important criterion is used for optimization, while the maximum permissible value is set for other criteria (i.e., all other criteria are transformed into MCO problem constraints) \cite{c39,c40}. Some researchers \cite{c41} describe the extended version of the $\varepsilon$-constraint method, in which violation of constraints is allowed if it leads to a noticeable improvement of values by certain criteria of an MCO problem. Application of this approach is rather complex because the maximum allowable values (the concessions) of criteria must be set initially before calculations begin, which often causes substantial difficulties. A proposed solution \cite{c42} of this issue is to combine the initial solution sequence for lexicographical optimization problems, when concessions can be set in the process of calculations after each subsequent optimization stage (although such a scheme again requires the solution of the series of scalar optimization problems). The approach proposed in \cite{c43} allows even greater freedom in formulating problems of searching for efficient decisions, when the order of criteria importance can also change dynamically. It is necessary to note that the presence of variability in problem statement for the search of efficient decisions is very important as requirements to optimality of selected decisions can be varied by the decision-maker in the course of calculations. Such variability is also provided in other approaches to solving MCO problems by various methods of criteria scalarization -- see, for example, \cite{c31,c44}. In \cite{c1} a combined approach is proposed to the application of scalarization criteria and additional constraints of the $\varepsilon$-constraint method.

@@

Without loss of generality, the values of the efficiency criteria are assumed to be non-negative, and their decrease corresponds to an increase in the effectiveness of the considered decisions $y \in D$. Problem (\ref{eq:1}) is considered in the context of the most complex decision-making problems, in which the efficiency criteria $f_i(y)$, $1 \leq i \leq s$, can be significantly multiextremal, and the procedures for calculating the values of the efficiency criteria at points of the search domain $y \in D$ can be computationally expensive. The criteria $f_i (y)$, $1 \leq i \leq s$ are also assumed to satisfy the Lipschitz condition
\begin{equation}\label{eq:3}
|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s,
\end{equation}
where $L_i$ is the Lipschitz constant for the function $f_i(y)$, $1 \leq i \leq s$, and ${\|*\|}$ denotes the Euclidean norm in $R^N$. The Lipschitz condition can be considered satisfied if for any variations of the decision $y \in D$ corresponding changes in the values of the criteria $f_i(y)$, $1 \leq i \leq s$, are limited.

@@

It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i \leq s$, the concessions may be quite rigid or, conversely, excessively large. In general, the change in the order of efficiency criteria importance may be required. Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:9})
\begin{equation}\label{eq:10}
%\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
\mathbb{F}=\{ F_1,F_2,\dots,F_t\},
\end{equation}
%new text begin
where $F_t, t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
dynamically by adding new required MCOlex problems or removing unnecessary ones. 
The problems of the set $\mathbb{F}$ specify various requirements for the desired solutions and, thus, all problems of the set $\mathbb{F}$ should be solved in the course of calculations.
In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex)\footnote{From a general point of view, the MCOlex problem can be considered as a special case of the MMCOlex problem for $t=1$.}.
%new text end

@@

For many years, global optimization has been the subject of many scientific studies. The results of research have been published in a large number of monographs -- see, for example, \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}. Making a reasonable choice of the approach for solving GO problems from the entire set of already existing global search algorithms requires taking into account the following.
\begin{itemize}
	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function at the point of the global minimum should be compared with the values of the function at all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain.
	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e. it must be dense only in the vicinity of globally optimal points.
	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}).
	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods. Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23}.
	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
\end{itemize}

@@

Dimension reduction by the Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27}. Numerical methods for constructing approximations of the Peano curves (evolvents) with a given accuracy are considered in \cite{c17,c18}.

@@

To provide uniform calculations, a sequential scheme is applied to obtain values the constraints $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$. According to this scheme, the constraint values are calculated strictly in their order, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme). The number of constraints, in which the values were calculated, will be referred below as the index $\nu=\nu(x)$, that is
\begin{equation}\label{eq:18}
\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j \leq \nu - 1, 1 \leq \nu = \nu(x) \leq m
\end{equation}
(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1})
\begin{equation}\label{eq:19}
F(x)=g_\nu(x), \nu=\nu(x),
\end{equation}
defined and computable everywhere in $[0,1]$. Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$). The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (further in the paper, this procedure will be referred to as trial).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) the function $F(x)$ from (\ref{eq:19}), b) the function $\Phi(x)$ from (\ref{eq:20})}
  \label{fig:1}
\end{figure}

@@

Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c18}. For instance, the following theorem is valid.

@@

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. 
%new text begin
This possibility of reusing search information can be briefly presented as follows.
%\begin{itemize}
%\item 
%При минимизации очередного критерия $f_l(y(x)), 1\leq l\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_l(x)$ запоминаются и вычисленные значения критериев $f_i, 1\leq i\leq k$.
%\item 
%При переходе к минимизации следующего критерия $f_{l+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений критериев $f_i, 1\leq i\leq k$ в поисковой информации $A_k$
%\begin{equation}\label{new_equation}
%z'_i=Z(f_i), 1\leq i\leq k,
%\end{equation}
%где процедура вычисления значений $z'_i, 1\leq i\leq k$ определяется правилами (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}) без повторных вычислениях значений критериев $f_l(y(x)), 1\leq l\leq s$.
%\end{itemize}
%new text end
This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.

Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCOlex problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
\begin{equation}\label{eq:30}
P_{\delta lex} (f,\delta,D)=Arg \min_{y \in D} {f_s(y)}, f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s,
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, the estimates $z_j^k$, ${1 \leq j < s}$, from (\ref{eq:23}) obtained from the search information $A_k$ can be used, and for the concessions $\delta_j$, $1 \leq j < s$, the normalized values can be applied, that is
\begin{equation}\label{eq:31}
	\begin{matrix}
	\delta_j = z_j^k + \theta_j (g_j^k-z_j^k ), 0 \leq \theta_j \leq 1, 1 \leq j < s,\\
	g_j^k = \max{\{ g_j (x_i ) : \nu(x_i)=j, 1 \leq i \leq k \}}.
	\end{matrix}
\end{equation}
%new text begin
%При этом, как и ранее, при любых изменениях оценок  $f_j^*, \delta_j, g_j^k, 1\leq j < s, 1 \leq i \leq k,$ решение задачи (\ref{eq:30}) может быть продолжено с использованием всей имеющейся поисковой информации после быстро выполняемых вычислений новых значений $z'_i,1\leq i \leq k,$ в соответствии с процедурой (\ref{new_equation}).
%new text end
In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.

@@

\begin{theorem}
Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differs from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
\begin{equation}\label{eq:32}
|\psi(x)-\phi(x)| \leq \delta, x \in [0,1],
\end{equation}
and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
\begin{equation}\label{eq:33}
|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
\end{equation}
if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
\begin{equation}\label{eq:34}
m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
\end{equation}
where
\begin{equation}\label{eq:35}
m=\max{r_l H_l^k}, 1 \leq l \leq M^k, k>k_0,
\end{equation}
\begin{equation}\label{eq:36}
L=\max{L_i}, 1 \leq i \leq s, L_i\; \text{from}\; (\ref{eq:3}),
\end{equation}
\begin{equation}\label{eq:37}
\psi(x^*)=\min{\psi(x)},x \in [0,1],
\end{equation}
\begin{equation}\label{eq:38}
x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k.
\end{equation}
\end{theorem}
The points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively.

@@

In this paper a new approach is proposed for solving computationally expensive lexicographic multicriteria optimization problems (MCOlex), which can be complicated by the multiextremal nature of the efficiency criteria and extensive volume of computations required to calculate the criteria values. A key feature of this class of problems is the ability to alter the ranking of efficiency criteria in terms of importance during calculations, which in turn necessitates a multistage solution for MCOlex problems.  This new type of problems is referred to as MMCOlex problems.

@@

\bibitem{c48} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information, J. Glob. Optim., \textbf{71}(1), 73--90 (2018). doi:10.1007/s10898-018-0624-3.

@@

\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly pareto compliant quality indicator. Math. Comput. Appl. \textbf{22}(1), 25 (2017)

@@

\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Digabel, S.Le, Salomon, L.: Performance indicators in multiobjective optimization. Technical Report G-2018-90, Les cahiers du GERAD. (2018)

@@

\bibitem{c59} Lai, L., Fiaschi, L., Cococcioni, M.: Solving mixed pareto-lexicographic multi-objective optimization problems: the case of priority chains. Swarm Evol. Comput. \textbf{55}, 100687 (2020) DOI: 10.1016/j.swevo.2020.100687

>>>>

\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}

\usepackage{xcolor}

@@

A noteworthy approach to solving MCO problems is lexicographic optimization methods. Such methods involve ranking the criteria by the degree of their importance and using the obtained ranking for subsequent optimization \cite{c3}. When criteria are ranked by importance, an MCO problem can be solved via consecutive optimization of criteria in the order of their importance: the most important criterion is optimized first, followed by the optimization of the second criterion based on the analysis of the set of decisions, in which the first criterion reaches the best value. This sequence is simple and clear and allows using many efficient scalar optimization algorithms. Such a multistage sequence can be reduced to solving one scalar optimization problem by linear convolution of criteria \cite{c38}. At the same time, this approach excessively reduces the number of calculated efficient decisions, which limits the number of options for the decision-maker. The $\varepsilon$-constraint scalarization approach can be used to expand the set of efficient decisions that can be considered by the decision-maker. In this case, only the most important criterion is used for optimization, while the maximum permissible value is set for other criteria (i.e., all other criteria are transformed into MCO problem constraints) \cite{c39,c40}. Some researchers \cite{c41} describe the extended version of the $\varepsilon$-constraint method, in which violation of constraints is allowed if it leads to a noticeable improvement of values by certain criteria of an MCO problem. Application of this approach is rather complex because the maximum allowable values (the concessions) of criteria must be set initially before calculations begin, which often causes substantial difficulties. A proposed solution \cite{c42} of this issue is to combine the initial solution sequence for lexicographical optimization problems, when concessions can be set in the process of calculations after each subsequent optimization stage (although such a scheme again requires the solution of the series of scalar optimization problems). The approach proposed in \cite{c43} allows even greater freedom in formulating problems of searching for efficient decisions, when the order of criteria importance can also change dynamically. It is necessary to note that the presence of variability in problem statement for the search of efficient decisions is very important as requirements to optimality of selected decisions can be varied by the decision-maker in the course of calculations. Such variability is also provided in other approaches to solving MCO problems by various methods of criteria scalarization -- see, for example, \cite{c31,c44}. In \cite{c1} a combined approach is proposed to the application of scalarization criteria and additional constraints of the $\varepsilon$-constraint method.

@@

Without loss of generality, the values of the efficiency criteria are assumed to be non-negative, and their decrease corresponds to an increase in the effectiveness of the considered decisions $y \in D$. Problem (\ref{eq:1}) is considered in the context of the most complex decision-making problems, in which the efficiency criteria $f_i(y)$, $1 \leq i \leq s$, can be significantly multiextremal, and the procedures for calculating the values of the efficiency criteria at points of the search domain $y \in D$ can be computationally expensive. The criteria $f_i (y)$, $1 \leq i \leq s$, are also assumed to satisfy the Lipschitz condition
\begin{equation}\label{eq:3}
|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s,
\end{equation}
where $L_i$ is the Lipschitz constant for the function $f_i(y)$, $1 \leq i \leq s$, and ${\|*\|}$ denotes the Euclidean norm in $R^N$. The Lipschitz condition can be considered satisfied if for any variations of the decision $y \in D$ corresponding changes in the values of the criteria $f_i(y)$, $1 \leq i \leq s$, are limited.

@@

It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i \leq s$, the concessions may be quite rigid or, conversely, excessively large. In general, the change in the order of efficiency criteria importance may be required. Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:9})
\begin{equation}\label{eq:10}
%\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
\mathbb{F}=\{ F_1,F_2,\dots,F_t\},
\end{equation}
{\color{red} 
where $F_t, t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
dynamically by adding new required MCOlex problems or removing unnecessary ones. 
The problems of the set $\mathbb{F}$ specify various requirements for the desired solutions and, thus, all problems of the set $\mathbb{F}$ should be solved in the course of calculations.
In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex)\footnote{From a general point of view, the MCOlex problem can be considered as a special case of the MMCOlex problem for $t=1$.}.
}

@@

For many years, global optimization has been the subject of many scientific studies. The results of research have been published in a large number of monographs -- see, for example, \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}. Making a reasonable choice of the approach for solving GO problems from the entire set of already existing global search algorithms requires taking into account the following.
\begin{itemize}
	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function at the point of the global minimum should be compared with the values of the function at all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain.
	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e. it must be dense only in the vicinity of globally optimal points.
	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}).
	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods. Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47,Grishagin2016_1,Grishagin2019}, the Peano space-filling curves \cite{c17,c23}.
	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
\end{itemize}

@@

Dimension reduction by the Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27,Grishagin2016_2}. Numerical methods for constructing approximations of the Peano curves (evolvents) with a given accuracy are considered in \cite{c17,c18}.

@@

To provide uniform calculations, a sequential scheme is applied to obtain values the constraints $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$. According to this scheme, the constraint values are calculated strictly in their order, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme). The number of constraints, in which the values were calculated, will be referred below as the index $\nu=\nu(x)$, that is
\begin{equation}\label{eq:18}
\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j \leq \nu - 1, 1 \leq \nu = \nu(x) \leq m
\end{equation}
(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1},
{\color{red}  the example taken from \cite{Book2013}})
\begin{equation}\label{eq:19}
F(x)=g_\nu(x), \nu=\nu(x),
\end{equation}
defined and computable everywhere in $[0,1]$. Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$). The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (further in the paper, this procedure will be referred to as trial).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) {\color{red} initial problem functions} , b) the function $F(x)$ from (\ref{eq:19})}
  \label{fig:1}
\end{figure}

@@

Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c18}.
{\color{red}
В частности, одна из модификаций связана с использованием информации о локальном поведении функций задачи \cite{Sergeyev2003}.
Условия сходимости алгоритма определяются следующей теоремой из \cite{c18}.
}

@@

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_j (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. 
{\color{red} 
This possibility of reusing search information can be briefly presented as follows.
\begin{itemize}
\item 
При минимизации очередного критерия $f_j(y(x)), 1\leq j\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_j(y(x))$ запоминаются и вычисленные значения векторного критерия $f(y(x_i)), 1\leq i\leq k$.
\item 
При переходе к минимизации следующего критерия $f_{j+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений векторного критерия $f(y(x_i)), 1\leq i\leq k,$ в поисковой информации $A_k$
\begin{equation}\label{new_equation}
z'_i=Z(f(y(x_i))), 1\leq i\leq k,
\end{equation}
где процедура вычисления значений $z'_i, 1\leq i\leq k$ определяется правилами (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}), без повторных вычислениях значений критериев $f_j(y(x)), 1\leq l\leq s$.
\end{itemize}
}
This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.

Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCOlex problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
\begin{equation}\label{eq:30}
P_{\delta lex} (f,\delta,D)=Arg \min_{y \in D} {f_s(y)}, f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s,
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, the estimates $z_j^k$, ${1 \leq j < s}$, from (\ref{eq:23}) obtained from the search information $A_k$ can be used, and for the concessions $\delta_j$, $1 \leq j < s$, the normalized values can be applied, that is
\begin{equation}\label{eq:31}
	\begin{matrix}
	\delta_j = z_j^k + \theta_j (g_j^k-z_j^k ), 0 \leq \theta_j \leq 1, 1 \leq j < s,\\
	g_j^k = \max{\{ g_j (x_i ) : \nu(x_i)=j, 1 \leq i \leq k \}}.
	\end{matrix}
\end{equation}
{\color{red} 
При этом, как и ранее, при любых изменениях оценок  $f_j^*, \delta_j, g_j^k, 1\leq j < s, 1 \leq i \leq k,$ решение задачи (\ref{eq:30}) может быть продолжено с использованием всей имеющейся поисковой информации после быстро выполняемых вычислений новых значений $z'_i,1\leq i \leq k,$ в соответствии с процедурой (\ref{new_equation}).
}
In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.

@@

\begin{theorem}
Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
\begin{equation}\label{eq:32}
|\psi(x)-\phi(x)| \leq \delta, x \in [0,1],
\end{equation}
and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
\begin{equation}\label{eq:33}
|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
\end{equation}
if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
\begin{equation}\label{eq:34}
m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
\end{equation}
where
\begin{equation}\label{eq:35}
m=\max{r_l H_l^k}, 1 \leq l \leq M^k, k>k_0,
\end{equation}
\begin{equation}\label{eq:36}
L=\max{L_i}, 1 \leq i \leq s, \; L_i\; \text{from}\; (\ref{eq:3}),
\end{equation}
\begin{equation}\label{eq:beta}
\beta =  \frac{2r\delta} {L\varepsilon(r-1)},
\end{equation}
\begin{equation}\label{eq:37}
\psi(x^*)=\min{\psi(x)},x \in [0,1],
\end{equation}
\begin{equation}\label{eq:38}
x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k.
\end{equation}
\end{theorem}
The points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively.

@@

In this paper a new approach is proposed for solving computationally expensive lexicographic multicriteria optimization problems (MCOlex), which can be complicated by the multiextremal nature of the efficiency criteria and extensive volume of computations required to calculate the criteria values. A key feature of this class of problems is the ability to alter the ranking of efficiency criteria in terms of importance during calculations, which in turn necessitates a multistage solution for MCOlex problems. {\color{red} This new type of problems is referred to as MMCOlex problems.}

@@

\bibitem{c48} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information, J. Glob. Optim., \textbf{71}(1), 73--90 (2018). DOI:10.1007/s10898-018-0624-3.

@@

\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly Pareto compliant quality indicator. Math. Comput. Appl. \textbf{22}(1), 25 (2017)

@@

\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Le Digabel, S., Salomon, L.: Performance indicators in multiobjective optimization. Technical Report G-2018-90, Les cahiers du GERAD. (2018)

@@

\bibitem{c59} Lai, L., Fiaschi, L., Cococcioni, M.: Solving mixed Pareto-lexicographic multi-objective optimization problems: the case of priority chains. Swarm Evol. Comput. \textbf{55}, 100687 (2020) DOI: 10.1016/j.swevo.2020.100687

\bibitem{Book2013} Strongin, R.G., Gergel, V.P., Grishagin, V.A., Barkalov, K.A.: Parallel computing in global optimization problems. MSU Press, Moscow. (2013, in Russian)

\bibitem{Sergeyev2003} Sergeyev, Y.D., Pugliese, P., Famularo, D.: Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints. Math. Program., Ser. A \textbf{96}, 489--512 (2003)

\bibitem{Grishagin2016_1} Grishagin, V.A., Israfilov, R.A.: Global search acceleration in the nested optimization scheme. In: AIP Conf. Proc., \textbf{1738}, 400010 (2016). DOI: 10.1063/1.4952198

\bibitem{Grishagin2016_2} Grishagin, V.A., Israfilov, R.A., Sergeyev, Y.D.: Comparative efficiency of dimensionality reduction schemes in global optimization. In: AIP Conf. Proc., \textbf{1776}, 060011 (2016). DOI: 10.1063/1.4965345

\bibitem{Grishagin2019} Gergel V.P., Grishagin V.A., Israfilov R.A. Adaptive dimensionality reduction in multiobjective optimization with multiextremal criteria. Lecture Notes in Computer Science, \textbf{11331}, 129--140 (2019)
